Hierarchically Supervised Latent Dirichlet Allocation

### 2011-05-31 paper/main.tex
Split inference section out.

\section{Inference}

\label{sec:inference} In the Bayesian approach to statistical modeling,
the primary task of inference is to find the posterior distribution
over the unobserved parameters of the model. However, it is often
possible and desirable to integrate over certain variables in the
model, also known as collapsing. In our model, it will often be the
case that the set of labels $\mathcal{L}$ is not fully observed for
every document. We will define $\mathcal{L}_{d}$ to be the subset
of labels which have been observed for document $d$. It is straightforward
to integrate out the variables $a_{l^{\prime},d}$ and $y_{l^{\prime},d}$
for $l^{\prime}\in\mathcal{L}\backslash\mathcal{L}_{d}$ from the
full generative model. We can also integrate out the parameters $\mathbf{\phi_{1:K}}$
and $\mathbf{\theta_{1:D}}$ as in \citet{Griffiths04}. Therefore,
in our model the latent variables are $\mathbf{z}=\{z_{1:N_{d},d}\}_{d=1,\ldots,D},\mathbf{\eta}=\{\mathbf{\eta_{l}}\}_{l\in\mathcal{L}},\mathbf{a}=\{a_{l^{\prime},d}\}_{l^{\prime}\in\mathcal{L}_{d},d=1,\ldots,D},\mathbf{\beta},\alpha,\alpha^{\prime}$
and $\gamma$.

The posterior distribution we seek cannot be solved in closed form.
This is often the case in evaluating posterior distributions of non-trivial
probabilistic models. We will appeal to one of the common methods
for approximating posterior distributions in the face of intractable
normalization factors: Markov chain Monte Carlo (MCMC) sampling. Since
in this model it is possible to sample from the conditional distributions
for all variables we will use the Gibbs sampling algorithm to obtain
an approximation to this posterior. %In particular, we will derive a collapsed Gibbs sampler.
%Given an observation of a set of observed labels and a document, the posterior distribution for the latent variables is
%\begin{equation}
%p\left(\theta,z_{1:N},\phi_{1:K},\eta_{i_{l,c}\in\mathcal{I}},a_{i_{l,c}\in\mathcal{I}},\beta,\alpha,\alpha',\gamma\mid w_{1:N},y_{i_{l,c}\in\mathcal{I}};\sigma,\lambda\right)\label{eq:Posterior}\end{equation}
%\[
%=\frac{p\left(\theta,z_{1:N},\phi_{1:K},\eta_{i_{l,c}\in\mathcal{I}},a_{i_{l,c}\in\mathcal{I}},\beta,\alpha,\alpha',\gamma,w_{1:N},y_{i_{l,c}\in\mathcal{I}};\sigma,\lambda\right)}{\int_{\theta,\phi,a,\eta,\alpha,\alpha',\beta,\gamma}\sum_{z}p\left(\theta,z_{1:N},\phi_{1:K},\eta_{i_{l,c}\in\mathcal{I}},a_{i_{l,c}\in\mathcal{I}},\beta,\alpha,\alpha',\gamma,w_{1:N},y_{i_{l,c}\in\mathcal{I}};\sigma,\lambda\right)}\]

\subsection{Gibbs Sampler}

We derive a collapsed Gibbs sampler for this model by considering
the individual conditional probability distributions for each of the
unobserved variables. We use the notation $\mathbf{z}_{-(n,d)}$ to
denote $\mathbf{z}\backslash z_{n,d}$. %, marginalizing
%$\theta_{1:D}$ and $\phi_{1:K}$. For details regarding collapsing
%in LDA models see \citet{Griffiths04}.

\subsubsection{$p\left(z_{n,d}\mid\mathbf{z_{-\left(n,d\right)}},\mathbf{a},\mathbf{w},\mathbf{\eta},\alpha,\beta,\gamma\right)$}

First we consider the conditional distribution of the assignment variable
for each word $n=1,\ldots,N_{d}$ in documents $d=1,\ldots,D$. The
conditional distribution does not include $\theta_{1:D}$ and $\phi_{1:K}$
because they have been integrated out as in the collapsed Gibbs sampler
\citep{Griffiths04}. The conditional distribution of $z_{n,d}$ is
proportional to the joint distribution of its markov blanket. %\begin{equation}
%p\left(z_{d,n}\mid\mathbf{z_{-\left(d,n\right)}},\mathbf{a},\mathbf{w},\mathbf{\eta},\alpha,\beta,\gamma\right)\propto p\left(z_{d,n},\mathbf{z_{-\left(d,n\right)}},\mathbf{a},\mathbf{w},\mathbf{\eta},\alpha,\beta,\gamma\right)\end{equation}
%Due to the factorization of this model, we can rewrite the joint distribution
%as the following:
\begin{equation}
p\left(z_{n,d}\mid\mathbf{z_{-\left(n,d\right)}},\mathbf{a},\mathbf{w},\mathbf{\eta},\alpha,\beta,\gamma\right)\propto\prod_{l\in\mathcal{L}_{d}}p\left(a_{l,d}\mid\mathbf{z},\eta_{l}\right)p\left(z_{n,d}\ |\ \mathbf{z_{-\left(n,d\right)}},\mathbf{a},\mathbf{w},\alpha,\beta,\gamma\right).\end{equation}
 The product is only over the subset of labels $\mathcal{L}_{d}$
which have been observed for document $d$. By isolating terms that
depend on $z_{n,d}$ and absorbing all other terms into a normalizing
constant as in \citep{Griffiths04} we find \begin{eqnarray}
\lefteqn{p\left(z_{n,d}=k\mid\mathbf{z_{-\left(n,d\right)}},\mathbf{a},\mathbf{w},\mathbf{\eta},\alpha,\beta,\gamma\right)\propto}\nonumber \\
 & \hspace{2cm}\left(n_{\left(\cdot\right),d}^{k,-\left(n,d\right)}+\alpha\beta_{k}\right)\frac{n_{w_{n,d},\left(\cdot\right)}^{k,-\left(n,d\right)}+\gamma}{\left(n_{\left(\cdot\right),\left(\cdot\right)}^{k,-\left(n,d\right)}+V\gamma\right)}\prod_{l\in\mathcal{L}_{d}}\exp\left\{ -\frac{\left(\bar{z}_{d}^{T}\eta_{l}-a_{l,d}\right)^{2}}{2}\right\} .\label{eq:z-likelihood}\end{eqnarray}
 Here, $n_{v,d}^{k,-\left(n,d^{\prime}\right)}$ represents the number
of words of type $v$ in document $d$ assigned to topic $k$ omitting
the $n^{{\rm th}}$ word of document $d^{\prime}$. The notation $(\cdot)$
in the subscript means the count resulting from summing over the omitted
subscript variable. Given Equation \ref{eq:z-likelihood}, $p\left(z_{d,n}\mid\mathbf{z_{-\left(d,n\right)}},\mathbf{a},\mathbf{w},\mathbf{\eta},\alpha,\mathbf{\beta},\gamma\right)$
can be sampled through enumeration.

\subsubsection{$p\left(\mathbf{\eta_{l}}\mid\mathbf{z},\mathbf{a},\sigma\right)$}

We now consider the conditional distribution of the regression coefficients
$\mathbf{\eta_{l}}$ for $l\in\mathcal{L}$. Given that $\mathbf{\eta_{l}}$
and $a_{l,d}$ are distributed normally, the posterior distribution
of $\mathbf{\eta_{l}}$ is normally distributed with mean $\hat{\mu}$
and covariance $\hat{\mathbf{\Sigma}}$ such that % (probably not the right place for this) We evaluated the model over various values of $\sigma$ where $\sigma=\left\{ 0.01,0.1,0.25,1,2\right\} $.
\begin{eqnarray}
\hat{\mathbf{\Sigma}}^{-1} & = & \mathbf{I}\sigma^{-1}+\bar{\mathbf{Z}}^{T}\bar{\mathbf{Z}}\\
\hat{\mu} & = & \hat{\mathbf{\Sigma}}_{i}\left(\mathbf{1}\frac{\mu}{\sigma}+\bar{\mathbf{Z}}^{T}\mathbf{a}_{l}\right).\end{eqnarray}
 This is a standard result from normal Bayesian linear regression
\citep{BishopPRML}. Here, $\bar{\mathbf{Z}}$ is a $D\times K$ matrix
such that row $d$ of $\mathbf{\bar{Z}}$ is $\bar{z}_{d}$, and $\mathbf{a}_{l}=[a_{l,1},a_{l,2},\ldots,a_{l,D}]^{T}$.
%p\left(\eta_{i_{l,c}}\mid\mathbf{z}_{1:D},\mathbf{a};\sigma\right)=\mathcal{N}\left(\eta_{i_{l,c}}\mid\hat{\mu}_{i},\hat{\Sigma}_{i}\right)\end{equation}
%\[
%\[

\subsubsection{$p\left(a_{l,d}\mid\mathbf{z},\mathbf{Y},\mathbf{\eta}\right)$}

%and \textmd{$p\left(y_{m,i}\mid\mathbf{a}\right)$}}
The auxiliary variables $a_{l,d}$ must be sampled for documents $d=1,\ldots,D$
and $l\in\mathcal{L}_{d}$. The conditional posterior distribution
of $a_{l,d}$ is the truncated normal distribution \begin{equation}
p\left(a_{l,d,}\mid\mathbf{z},\mathbf{Y},\mathbf{\eta}\right)\propto\frac{1}{\sqrt{2\pi}}exp\left\{ -\frac{1}{2}\left(a_{l,d}-\mathbf{\eta_{l}}^{T}\mathbf{\bar{z}}_{d}\right)\right\} I\left(a_{l,d}y_{l,d}>0\right).\end{equation}
 This conditional distribution can be sampled using an inverse CDF
method. %However, if $y_{d,i_{l,c}}$ is unobserved then $a_{d,i_{l,c}}$must
%be sampled jointly with $y_{d,i_{l,c}}$ to ensure that the Markov
%chain is ergodic. Suppose that $a_{d,i_{l,c}}$ is sampled to have
%a negative value and $y_{d,i_{l,c}}$ is apporopriately sampled at
%-1. Although there exist valid states where $a_{d,i_{l,c}}>0$ and
%$y_{d,i_{l,c}}=1$, they will never be reached by such a Markov chain
%since $p\left(a_{d,i_{l,c}}<0\mid y_{d,i_{l,c}}=-1\right)=1$ and
%$p\left(y_{d,i_{l,c}}=-1\mid a_{d,i_{l,c}}<0\right)=1$. Therefore,
%to ensure ergodicity, $a_{d,i_{l,c}}$and $y_{d,i_{l,c}}$ must be
%sampled from the joint distribution as shown in Equation \ref{eq:Probit-Joint}.
%\begin{equation}
%p\left(a_{d,i_{l,c}},y_{d,i_{l,c}}\mid\mathbf{z},\mathbf{Y}_{-\left(d,i_{l,c}\right)},\mathbf{\eta}\right)\propto p\left(y_{i_{l,c}}\mid\mathbf{a},\mathbf{y}_{-\left(l,c\right)}\right)p\left(a_{d,i_{l,c}}\mid\mathbf{z},\mathbf{Y},\mathbf{\eta}\right)\end{equation}
%\begin{equation}
%p\left(y_{i_{l,c}}\mid\mathbf{a},\mathbf{y}_{-\left(l,c\right)}\right)=\delta\left(sign\left(a_{d,i_{l,c}}\right)=y_{i_{l,c}}\right)p\left(y_{i_{l,c}}\mid y_{parents_{l,c}}\right)\prod_{i_{\hat{l},\hat{c}}\in children_{l,c}}p\left(y_{i_{\hat{l},\hat{c}}}\mid y_{i_{l,c}}\right)\end{equation}
%\begin{equation}
%p\left(y_{i_{l,c}}=-1\mid y_{parent{}_{l,c}}\right)=\begin{cases}
%1, & y_{parent_{l,c}}=-1\\
%0.5, & y_{parent_{l,c}}=1\end{cases}\end{equation}
%\[
%p\left(a_{d,i_{l,c}},y_{d,i_{l,c}}\mid\mathbf{z},\mathbf{Y}_{-\left(d,i_{l,c}\right)},\mathbf{\eta}\right)\]
%\begin{equation}
%=\begin{cases}
%\mathcal{N}\left(a_{d,i_{l,c}}\mid\bar{z}^{T}\eta_{i_{l,c}},1\right)p\left(y_{d,i_{l,c}}\mid a_{d,i_{l,c}}\right), & y_{parent_{l,c}}=1,\forall y_{i_{\hat{l},\hat{c}}}\in y_{children_{l,c}},y_{i_{\hat{l},\hat{c}}}=-1\\
%trunc\mathcal{N}^{-}\left(a_{d,i_{l,c}}\mid\bar{z}^{T}\eta_{i_{l,c}},1\right)\delta\left(y_{d,i_{l,c}}=-1\right), & y_{parent_{l,c}}=-1\\
%trunc\mathcal{N}^{+}\left(a_{d,i_{l,c}}\mid\bar{z}^{T}\eta_{i_{l,c}},1\right)\delta\left(y_{d,i_{l,c}}=1\right), & \exists y_{i_{\hat{l},\hat{c}}}\in y_{children_{l,c}}\setminus y_{i_{\hat{l},\hat{c}}}=1\\
%0 & otherwise\end{cases}\end{equation}
%where $\mathbf{Y}_{-\left(d,i_{l,c}\right)}$ denotes all of the response
%variables excluding the response variable being sampled.

\subsubsection{$p\left(\beta\mid\mathbf{z},\alpha^{\prime},\alpha\right)$}

In our model, we place a hierarchical Dirichlet prior over topic assignments.
This flexible distribution allows for an asymmetric prior over document
level distributions over topics \citep{WallachMiMc2009}. This prior
shares many features with the hierarchical Dirichlet process and inference
over this distribution proceeds in a very similar fashion.

Posterior inference is performed using the {}``direct assignment''
method of \citet{TehJorBea2006}. \begin{equation}
\beta\sim{\rm Dir}\left(m_{\left(\cdot\right),1}+\alpha^{\prime},m_{\left(\cdot\right),2}+\alpha^{\prime},\ldots,m_{\left(\cdot\right),K}+\alpha^{\prime}\right)\end{equation}
 \begin{equation}
p\left(m_{d,k}=m\mid\mathbf{z},\mathbf{m}_{-\left(d,k\right)},\beta\right)=\frac{\Gamma\left(\alpha\beta_{k}\right)}{\Gamma\left(\alpha\beta_{k}+n_{d,k}\right)}s\left(n_{d,k},m\right)\left(\alpha\beta_{k}\right)^{m}\end{equation}
 where $s\left(n,m\right)$ represents stirling numbers of the first
kind.

\subsubsection{$p\left(\alpha\right)$, $p\left(\alpha'\right)$, $p\left(\gamma\right)$}

The hyperparameters $\alpha$, $\alpha^{\prime}$, and $\gamma$ are
given broad ${\rm Gamma}(1,1000)$ prior distributions and sampled
via the Metropolis-Hastings algorithm.

\section{Experiments}

>>>>

\section{Inference}
\label{sec:inference}
\input{inference}
\section{Experiments}
